The Heavy Photon Search Test Detector

### 2013-12-04 hps-testrun-nim.tex
Changed EoverP plot to Pelles based on good region. It has loose track selection and the calibration is the default as of v5 detector.
Added new t0 plot. Fixed biblio style. Changed to a test run ecal beamline picture.

%% main text
\section{Introduction}
\label{introduction}
One of the more striking anomalies in recent astrophysical observations is an excess of high energy 
electrons and positrons consistent with a large, unknown source of pair 
production~\cite{Adriani:2008zr, FermiLAT:2011ab}. The existence of a massive, hidden sector vector 
boson; a ``dark photon;" that mixes kinetically with the Standard Model photon and mediates 
annihilation or decay of dark matter particles is a theoretically attractive explanation of this cosmic-ray 
excess~\cite{Holdom:1985ag}\cite{ArkaniHamed:2008qn}.
\begin{center}
{\small
\begin{figure*}[ht]
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test apparatus installed on the beam line. SHOULD FIND A PIC WITH LOWER HEIGHT.}
\label{fig:testrundetector}
\end{figure*}
}
\end{center}

@@

\subsection{Digitization, Event Building and Data Transmission}
\label{sec:svt_daq}
The SVT data acquisition (DAQ) based on the Reconfigurable Cluster Element (RCE) and Cluster 
Interconnect (CI) concept developed at SLAC as generic building blocks for DAQ systems in high energy 
physics {\bf [need reference]}. 
The RCE is a generic computational building block, housed on a separate daughter card called 
Data Processing Module (DPM), that are realized on an ATCA front board called the Cluster On Board (COB), 
see Fig.~\ref{fig:cob}.
 \begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/svt_daq_module}
	\caption{Picture of the SVT DAQ COB board with four data processing daughter cards visible.}
	\label{fig:cob}
}
\end{center}
\end{figure}
The Gen-I RCE used in the Test run consisted of a Virtex~5 FPGA with 1~GB of DDR3 RAM. 
The outputs of 12 SVT half-modules are digitized on the Rear-Transition-Module (RTM), a custom board on the back side of the ATCA crate, interfacing the HPS readout to the generic DAQ components on the COB. 
A pre-amplifier converts the APV25 differential current output to a different voltage output scaled to the sensitive 
range of a 14-bit ADC operating at the system clock of 41.667~MHz. 
The RTM is organized into four sections with each section supporting three SVT half-module hybrids (15 
APV25 channels). The RTM also includes a 4-channel Fiber Optic module and supporting logic which can be 
used to interface to the JLab trigger system supervisor.
Each section of the RTM, up to three hybrids, is input to a DPM which apply thresholds for data reduction and 
organizes the sample data into UDP datagrams. The DPM also hosts an I2C controller used to configure and 
monitor the APV25 chips on those three hybrids. A single ATCA crate with two COB cards was used, one 
supporting four DPMs and one supporting 3 DPMs and one DPM that is configured as the trigger and data 
transmission module. The two COB cards and their DPMs are interconnected with a 10Gb/s switch 
card~\cite{Larsen:2011zb} which 
also hosts two 1Gb/s Ethernet interfaces to the external SVT DAQ PC.

@@

%Step\subsubsection{PbWO$_4$ Crystals}
The HPS test run ECal uses PbWO$_{4}$ crystals, APDs, and preamplifier boards from the inner calorimeter (IC) of the CLAS detector\cite{2008erhm.conf..421N}.  In Figure~\ref{fig:ecal-module} a schematic view of the module assembly is shown. Lead-tungstate crystals are $160$ mm long and are tapered with front face of $13.3\times13.3$~mm$^2$, tapered to $16\times16$~mm$^2$ at the rear face. The crystals were fabricated with very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal is wrapped in VM2000 multilayer polymer mirror film.
The APDs (Hamamatsu S8664-55) with $5\times5$~mm$^2$ active area and $75\%$ quantum efficiency were glued on the rear face of the crystals using MeltMount 1.7 thermal plastic adhesive. The low gain of APDs ($\sim 200$) was compensated with custom made preamplifier boards, which provide factor 2000 amplification of the ADP signal.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}

@@

During the test run ECal was mounted downstream of the analyzing dipole magnet at the distance of about $147$~cm from the 
upstream edge of the magnet. 
%The two ECal modules are positioned just above and below the ECal vacuum chamber as shown in bottom panel of Fig.~\ref{fig:ecal}. 
%The primary electron beam, scattered electrons and bremsstrahlung photons pass through the system in vacuum. 
The gap between innermost edge of the crystals of two parts was $7.4$~cm, centered on the beam plane. The temperature stability of the system during the test run was better than $1^\circ~$F, the temperature uniformity inside the box was better than  $4^\circ~$F.

\begin{center}
{\small
\begin{figure}
\includegraphics[width=7cm]{figures/ECal}
\includegraphics[width=7cm]{figures/ecal_vacuumchamber.pdf}
%\includegraphics[width=7cm]{figures/ecal_assembly}
\caption{Layout of PbWO$_{4}$ modules in two halves of the ECal (top) and ECal assembly together with ECal vacuum chamber (bottom). }
\label{fig:ecal}
\end{figure}
}
\end{center}
%Step \subsubsection{Support Structure}
%Step In order to stabilize the calorimeter's performance, the crystals, APDs, and amplifiers are enclosed within a temperature controlled environment, held constant at the level of .
%Step\subsubsection{Avalanche Photodiode Readout and Preamplifiers}
%Step\label{sec:apd}
%StepAvalanche Photodiodes (APD) S8664-55 produced by Hamamatsu Corp. are used as photodetectors. They have a $5\times5$~mm$^2$ active area and quantum efficiency of about 75\%. Single APDs were centered on the back ends of the crystals using MeltMount 1.7 thermal plastic adhesive. The low? amplitude signals of APDs are amplified using custom?made preamplifier boards. 
%The gain uniformity of the APDs is on the order of XXXX\%; with APD high voltage bias varying by as much as 100V. Since only a limited number of high voltage channels were available during the test, 10 channels were grouped together with an attempt to group modules with similar gains.

@@

ECal signal readout was organized as follows: the amplified APD signal from the motherboard is sent to a $2:1$ signal splitter. After the split, 2/3 of the signal is sent a discriminator then to a Time-to-Digital 
Converter (TDC). The 1/3 of the signal is fed to a single channel of 16-channel JLAB Flash Amplitude-to-Digital Converter (FADC250), see Fig.\ref{fig:fadc} \cite{fadc}.  
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Two 20-slot VXS crates with 14 (for top ECal) and 13 (for bottom ECal) FADC bards were employed for the test run.

The FADCs store 12-bit digitized samples at 250~MHz in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed for the readout. HPS used FADCs in the integration mode. A predefined number of samples of the signal that exceeds a predefined threshold have been summed and stored in 17-bit register for readout. The number of samples for integration was defined by two parameters that could be set separately, the number of samples integrated before the signal crossed threshold ($NSB$) and  the number of samples integrated after the signal crossed threshold ($NSA$), see Fig.\ref{fig:hps_trigger_data}. 
This scheme significantly compresses the data input 
to the FADC. During data analysis, a pedestal value is subtracted to obtain the actual summed energy.

For various reasons only 385 out of 442 modules (87\%) were operational during the test run. The readout threshold was set to $70$ MeV, the number of samples before and after the threshold crossing were NSB=5 and NSA=30, respectively.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=8cm]{figures/fadc_datapath_diagram.png}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}

@@

The HPS test run trigger system is designed to select \ee{} events efficiently by 
using information from the ECal. The trigger looks for time coincidences of clusters in the top and bottom 
half of the ECal. The trigger system can be broken down into three sections (see 
Fig.~\ref{fig:hps_trigger_cal}):
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. 
Sends the measured pulse energy and arrival time to the Crate Trigger Processor (CTP),
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. 
The cluster energy and arrival time are sent to the Sub-System Processor (SSP),
 \item SSP (cluster pair finding): searches for time coincidences of pairs of clusters from 
the top and bottom half of the ECal and applies topological selections.
 \end{itemize}
 \begin{figure}[t]
\begin{center}
{\small
 \includegraphics[width=8cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes 
signals for each detector channel and sends them for cluster finding in the CTP. The CTP clusters are 
sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the 
ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout 
the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
The time coincidence window of pairs of clusters in the top and bottom half of the ECal are 
programmable with 4~ns resolution.

As described above, the first stage of the trigger logic is incorporated into the 
FPGA's on the FADC boards. The trigger  
integrates using a "time-over-threshold" window, see Fig. \ref{fig:hps_trigger_data}.
For triggering purposes, every channel has: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item a trigger threshold, measured in ADC counts, 
 \item an offset, 
 \item a conversion factor to convert ADC counts into energy in MeV, from $50$ to $\sim1500$~MeV, with 5-bit accuracy 
 \item an energy threshold discriminator to provide aminimum energy cutoff
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The offset is subtracted from ADC samples before they are integrated.  
The values reported to the CTP are the 5-bit pulse energy and the 3-bit time at which the 
pulse crossed the threshold. Data from every channel is sent to the CTP every 32~ns. With the available 
3-bit time information, the CTP looks for time coincidences of crystal signals within an 8~ns window (the 
limit is 4~ns).
The cluster finding algorithm is fast and makes use of the parallel processing capability of the FPGA's by 
simultaneously searching for 125 clusters, up to 3x3 in size, across the calorimeter crystal array, and it 
performs the following tasks:
\begin{itemize}
\item Adds together the energy for every 3x3 square of channels in the ECal
\item Hits are added together if their leading edges occurred within a programmable number of 4~ns 
clock cycles  (HPS test run used 8~ns coincidence time interval).
\item If the 3x3 energy sum is larger than the programmable cluster energy threshold and the sum is 
greater than any neighboring 3x3 windows, the CTP reports the cluster parameters to the SSP. 
\end{itemize}
The CTP evaluates all the hits in its half of the calorimeter every 4~ns. A programmable time window is 
used to allow hits that are slightly out of time with each other to be considered as part of a cluster sum. 
This is done by reporting hits when they occur and then reporting them again for the next $N$ number of 
4~ns clock cycles, where $N \in [0,7]$. This is useful to deal with skew and jitter that develop from the 
detector, cabling, and electronics. As described above, the CTP only selects the 3x3 window with the 
highest energy sum of its neighbors. This filtering is applied to deal with overlapping clusters and cases 
where the cluster is larger than a 3x3 window.

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svt_daq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_110513_samples_L1_top.pdf}
	\caption{The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}

These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_mip_small.pdf}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit 
time with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter 
due to clock phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/t0_resolution_run1351_110513.pdf}
	\caption{
	The cluster time residual for a representative sensor relative to the track time (see text for more 
	details). \color{red}{\bf Fix axis.} }
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is somewhat 
worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate 
with the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

While occupancy was slightly larger than expected, good agreement between data and simulation was 
found after taking into account dead or noisy channels. The hit efficiency was estimated by measuring 
the number of good tracks with a hit close to the extrapolated intersection of a given sensor that was 
excluded from the track fit itself. Tracks which intersect regions with known bad channels or very close to 
in the edge region are excluded in the calculation of the single hit efficiency. The hit efficiency, see 
Fig.~\ref{fig:hit_efficiency}, was measured to be above 98\% and fairly uniform across the SVT.
\begin{figure}[h]
\begin{center}
{\small
    	\includegraphics[width=7cmh]{figures/single_hit_efficiency_Omar_11192013.pdf}
        \caption{ The hit reconstruction efficiency as a function of detector layer.}
	\label{fig:hit_efficiency}
}
\end{center}
\end{figure}
The spatial resolution of similar microstrip sensors is well established by test beam data, against which 
the charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise (S/N) and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ratio above 20, as demonstrated 
here, is relatively constant at approximately 6~$\mu$m for tracks that enters approximately normal to the 
sensors as in HPS.

@@

By selecting \ee{} pairs from the triggered events we are able to study basic distributions of pair 
production kinematics and in particular those related to our vertex performance. Pairs of opposite charge 
tracks, one in the top and one in the bottom half of the SVT, with momentum larger than 400~MeV were 
selected. The pair production kinematics are relatively well reproduced given the alignment of the 
tracker; Fig.~\ref{fig:pair_kin} shows the invariant mass and ratio of electron momentum over the sum of 
electron and positron. 
 \begin{center}
{\small
\begin{figure*}[ht]
   \includegraphics[width=5cm]{figures/h_trk_top_px_h_trk_top_px_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_invM_h_invM_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_sumE_h_sumE_trigsel4hit_pair1351_twotrkfilt-v6-paper}
\caption{Kinematic distributions for e$^+$e$^-$ pairs selected by opposite charged tracks in the top and bottom half of the tracker: track momentum in the top half of the SVT (left), invariant mass (middle) and the sum of the track momentum for the pair (right).}
\label{fig:pair_kin}
\end{figure*}
}
\end{center}
We estimate the accuracy of the SVT momentum scale and resolution by looking at the agreement 
between data and simulation in the shape of the kinematic distributions for single- and two track 
events.  By adjusting the overall momentum scale and smearing the simulated events with a 
Gaussian resolution we estimate an uncertainty of 10\% on both the momentum scale and 
resolution in the Test run. The expected momentum resolution from simulation is between 4-5\% 
for tracks in the momentum range of the Test run.

>>>>

%% main text
\section{Introduction}
\label{introduction}
One of the more striking anomalies in recent astrophysical observations is an excess of high energy 
electrons and positrons consistent with a large, unknown source of pair 
production~\cite{Adriani:2008zr, FermiLAT:2011ab}. The existence of a massive, hidden sector vector 
boson; a ``dark photon;" that mixes kinetically with the Standard Model photon and mediates 
annihilation or decay of dark matter particles is a theoretically attractive explanation of this cosmic-ray 
excess~\cite{Holdom:1985ag}\cite{ArkaniHamed:2008qn}.
\begin{center}
{\small
\begin{figure*}[ht]
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test apparatus installed on the beam line.}
\label{fig:testrundetector}
\end{figure*}
}
\end{center}

@@

\subsection{Digitization, Event Building and Data Transmission}
\label{sec:svt_daq}
The SVT data acquisition (DAQ) based on the Reconfigurable Cluster Element (RCE) and Cluster 
Interconnect (CI) concept developed at SLAC as generic building blocks for DAQ systems in high energy 
physics. The RCE is a generic computational building block, housed on a separate daughter card called 
Data Processing Module (DPM), that are realized on an ATCA front board called the Cluster On Board 
(COB), see Fig.~\ref{fig:cob}.
 \begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/svt_daq_module}
	\caption{Picture of the SVT DAQ COB board with four data processing daughter cards visible.}
	\label{fig:cob}
}
\end{center}
\end{figure}
The Gen-I RCE used in the Test run consisted of a Virtex~5 FPGA with 1~GB of DDR3 RAM. 
The outputs of 12 SVT half-modules are digitized on the Rear-Transition-Module (RTM), a custom board on the back side of the ATCA crate, interfacing the HPS readout to the generic DAQ components on the COB. 
A pre-amplifier converts the APV25 differential current output to a different voltage output scaled to the sensitive 
range of a 14-bit ADC operating at the system clock of 41.667~MHz. 
The RTM is organized into four sections with each section supporting three SVT half-module hybrids (15 
APV25 channels). The RTM also includes a 4-channel Fiber Optic module and supporting logic which can be 
used to interface to the JLab trigger system supervisor.
Each section of the RTM, up to three hybrids, is input to a DPM which apply thresholds for data reduction and 
organizes the sample data into UDP datagrams. The DPM also hosts an I2C controller used to configure and 
monitor the APV25 chips on those three hybrids. A single ATCA crate with two COB cards was used, one 
supporting four DPMs and one supporting 3 DPMs and one DPM that is configured as the trigger and data 
transmission module. The two COB cards and their DPMs are interconnected with a 10Gb/s switch 
card~\cite{Larsen:2011zb} which 
also hosts two 1Gb/s Ethernet interfaces to the external SVT DAQ PC.

@@

%Step\subsubsection{PbWO$_4$ Crystals}
The HPS test run ECal uses PbWO$_{4}$ crystals, APDs, and preamplifier boards from the inner calorimeter (IC) of the CLAS detector\cite{2008erhm.conf..421N}.  In Figure~\ref{fig:ecal-module} a schematic view of the module assembly is shown. Lead-tungstate crystals are $160$ mm long and are tapered with front face of $13.3\times13.3$~mm$^2$, tapered to $16\times16$~mm$^2$ at the rear face. The crystals were fabricated with very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal is wrapped in VM2000 multilayer polymer mirror film.
The APDs (Hamamatsu S8664-55) with $5\times5$~mm$^2$ active area and $75\%$ quantum efficiency were glued on the rear face of the crystals using MeltMount 1.7 thermal plastic adhesive. The low gain of APDs ($\sim 200$) was compensated with custom made preamplifier boards, which provide factor 2000 amplification of the ADP signal. 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}

@@

During the test run ECal was mounted downstream of the analyzing dipole magnet at the distance of about $147$~cm from the 
upstream edge of the magnet. 
%The two ECal modules are positioned just above and below the ECal vacuum chamber as shown in bottom panel of Fig.~\ref{fig:ecal}. 
%The primary electron beam, scattered electrons and bremsstrahlung photons pass through the system in vacuum. 
The gap between innermost edge of the crystals of two parts was $7.4$~cm, centered on the beam plane. The temperature stability of the system during the test run was better than $1^\circ~$F, the temperature uniformity inside the box was better than  $4^\circ~$F. 
\begin{center}
{\small
\begin{figure*}[htb]
\includegraphics[width=0.45\textwidth]{figures/ECal}
\includegraphics[width=0.45\textwidth]{figures/ecal_picture_beamline}
%\includegraphics[width=7cm]{figures/ecal_vacuumchamber.pdf}
%\includegraphics[width=7cm]{figures/ecal_assembly}
\caption{Layout of PbWO$_{4}$ modules in two halves of the ECal (left) and the ECal assembly on the beam line (right). }
\label{fig:ecal}
\end{figure*}
}
\end{center}
%Step \subsubsection{Support Structure}
%Step In order to stabilize the calorimeter's performance, the crystals, APDs, and amplifiers are enclosed within a temperature controlled environment, held constant at the level of .
%Step\subsubsection{Avalanche Photodiode Readout and Preamplifiers}
%Step\label{sec:apd}
%StepAvalanche Photodiodes (APD) S8664-55 produced by Hamamatsu Corp. are used as photodetectors. They have a $5\times5$~mm$^2$ active area and quantum efficiency of about 75\%. Single APDs were centered on the back ends of the crystals using MeltMount 1.7 thermal plastic adhesive. The low? amplitude signals of APDs are amplified using custom?made preamplifier boards. 
%The gain uniformity of the APDs is on the order of XXXX\%; with APD high voltage bias varying by as much as 100V. Since only a limited number of high voltage channels were available during the test, 10 channels were grouped together with an attempt to group modules with similar gains.

@@

ECal signal readout was organized as follows: the amplified APD signal from the motherboard is sent to a $2:1$ signal splitter. After the split, 2/3 of the signal is sent a discriminator then to a Time-to-Digital 
Converter (TDC). The 1/3 of the signal is fed to a single channel of 16-channel JLAB Flash Amplitude-to-Digital Converter (FADC250), see Fig.\ref{fig:fadc} \cite{fadc}.  
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=5cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Two 20-slot VXS crates with 14 (for top ECal) and 13 (for bottom ECal) FADC bards were employed for the test run.

The FADCs store 12-bit digitized samples at 250~MHz in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed for the readout. HPS used 
FADCs in the integration mode. A predefined number of samples of the signal that exceeds a predefined 
threshold have been summed and stored in 17-bit register for readout. The number of samples for 
integration was defined by two parameters that could be set separately, the number of samples 
integrated before the signal crossed threshold ($NSB$) and  the number of samples integrated after the 
signal crossed threshold ($NSA$), see Fig.\ref{fig:hps_trigger_data}. 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/fadc_datapath_diagram.png}
\caption{\small{FADC data paths.}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}
This scheme significantly compresses the data input to the FADC. During data analysis, a pedestal 
value is subtracted to obtain the actual summed energy.

For various reasons only 385 out of 442 modules (87\%) were operational during the test run. The 
readout threshold was set to $70$ MeV, the number of samples before and after the threshold crossing 
were 5 and 30, respectively.

@@

The HPS test run trigger system is designed to select \ee{} events efficiently by 
using information from the ECal. The trigger looks for time coincidences of clusters in the top and bottom 
half of the ECal. The trigger system can be broken down into three sections as shown in 
Fig.~\ref{fig:hps_trigger_cal}:
 \begin{figure}[b]
\begin{center}
{\small
 \includegraphics[width=8cm]{figures/hps_trigger_cal}
\caption{Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes 
signals for each detector channel and sends them for cluster finding in the CTP. The CTP clusters are 
sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the 
ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to readout 
the sub-detectors.}
 \label{fig:hps_trigger_cal}
}
\end{center}
 \end{figure}
 \begin{itemize}
 \item FADC (pulse finding): samples and digitizes the signal pulses from each detector channel. 
Sends the measured pulse energy and arrival time to the Crate Trigger Processor (CTP),
\item CTP (cluster finding): groups FADC pulses from each half of the ECal into clusters. 
The cluster energy and arrival time are sent to the Sub-System Processor (SSP),
 \item SSP (cluster pair finding): searches for time coincidences of pairs of clusters from 
the top and bottom half of the ECal and applies topological selections.
 \end{itemize}

The time coincidence window of pairs of clusters in the top and bottom half of the ECal are 
programmable with 4~ns resolution.  As described above, the first stage of the trigger logic is 
incorporated into the FPGA's on the FADC boards. The trigger  
integrates using a "time-over-threshold" window as explained in Fig. \ref{fig:hps_trigger_data}.
For triggering purposes, every channel has: 
\begin{itemize}
% \item number of samples integrated before the threshold crossing (NSB),
 %\item number of samples integrated after the  threshold crossing (NSA),
 \item a trigger threshold, measured in ADC counts, 
 \item an offset, 
 \item a conversion factor to convert ADC counts into energy in MeV, from $50$ to $\sim1500$~MeV, with 5-bit accuracy 
 \item an energy threshold discriminator to provide aminimum energy cutoff
 \end{itemize}
Note that the threshold for the trigger path can be set independently from the readout threshold.
The offset is subtracted from ADC samples before they are integrated.  
The values reported to the CTP are the 5-bit pulse energy and the 3-bit time at which the 
pulse crossed the threshold. Data from every channel is sent to the CTP every 32~ns. With the available 
3-bit time information, the CTP looks for time coincidences of crystal signals within an 8~ns window (the 
limit is 4~ns).
The cluster finding algorithm is fast and makes use of the parallel processing capability of the FPGA's by 
simultaneously searching for 125 clusters, up to 3x3 in size, across the calorimeter crystal array, and it 
performs the following tasks:
\begin{itemize}
\item Adds together the energy for every 3x3 square of channels in the ECal
\item Hits are added together if their leading edges occurred within a programmable number of 4~ns 
clock cycles  (HPS test run used 8~ns coincidence time interval).
\item If the 3x3 energy sum is larger than the programmable cluster energy threshold and the sum is 
greater than any neighboring 3x3 windows, the CTP reports the cluster parameters to the SSP. 
\end{itemize}
The CTP evaluates all the hits in its half of the calorimeter every 4~ns. A programmable time window is 
used to allow hits that are slightly out of time with each other to be considered as part of a cluster sum. 
This is done by reporting hits when they occur and then reporting them again for the next $N$ number of 
4~ns clock cycles, where $N \in [0,7]$. This is useful to deal with skew and jitter that develop from the 
detector, cabling, and electronics. As described above, the CTP only selects the 3x3 window with the 
highest energy sum of its neighbors. This filtering is applied to deal with overlapping clusters and cases 
where the cluster is larger than a 3x3 window.

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svt_daq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_110513_samples_L1_top.pdf}
	\caption{The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}

These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/run1351_mip_small.pdf}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit 
time with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter 
due to clock phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/test_run_1351_hit_time_corrected_top_layer2.pdf}
	\caption{
	The cluster time residual for a representative sensor relative to the track time (see text for more 
	details). }
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is somewhat 
worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate 
with the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

While occupancy was slightly larger than expected, good agreement between data and simulation was 
found after taking into account dead or noisy channels. The hit efficiency was estimated by measuring 
the number of good tracks with a hit close to the extrapolated intersection of a given sensor that was 
excluded from the track fit itself. Tracks which intersect regions with known bad channels or very close to 
in the edge region are excluded in the calculation of the single hit efficiency. The hit efficiency, see 
Fig.~\ref{fig:hit_efficiency}, was measured to be above 98\% and fairly uniform across the SVT.
\begin{figure}[]
\begin{center}
{\small
    	\includegraphics[width=7cm]{figures/single_hit_efficiency_Omar_11192013.pdf}
        \caption{ The hit reconstruction efficiency as a function of detector layer.}
	\label{fig:hit_efficiency}
}
\end{center}
\end{figure}
The spatial resolution of similar microstrip sensors is well established by test beam data, against which 
the charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise (S/N) and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ratio above 20, as demonstrated 
here, is relatively constant at approximately 6~$\mu$m for tracks that enters approximately normal to the 
sensors as in HPS.

@@

By selecting \ee{} pairs from the triggered events we are able to study basic distributions of pair 
production kinematics and in particular those related to our vertex performance. Pairs of opposite charge 
tracks, one in the top and one in the bottom half of the SVT, with momentum larger than 400~MeV were 
selected. The pair production kinematics are relatively well reproduced given the alignment of the 
tracker; Fig.~\ref{fig:pair_kin} shows the invariant mass and ratio of electron momentum over the sum of 
electron and positron. 
 \begin{center}
{\small
\begin{figure*}[t]
   \includegraphics[width=5cm]{figures/h_trk_top_px_h_trk_top_px_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_invM_h_invM_trigsel4hit_pair1351_twotrkfilt-v6-paper}
   \includegraphics[width=5cm]{figures/h_sumE_h_sumE_trigsel4hit_pair1351_twotrkfilt-v6-paper}
\caption{Kinematic distributions for e$^+$e$^-$ pairs selected by opposite charged tracks in the top and bottom half of the tracker: track momentum in the top half of the SVT (left), invariant mass (middle) and the sum of the track momentum for the pair (right).}
\label{fig:pair_kin}
\end{figure*}
}
\end{center}
We estimate the accuracy of the SVT momentum scale and resolution by looking at the agreement 
between data and simulation in the shape of the kinematic distributions for single- and two track 
events.  By adjusting the overall momentum scale and smearing the simulated events with a 
Gaussian resolution we estimate an uncertainty of 10\% on both the momentum scale and 
resolution in the Test run. The expected momentum resolution from simulation is between 4-5\% 
for tracks in the momentum range of the Test run.
